---
title: "STAT 3675Q Homework 1"
subtitle: "Due date: **Thursday, September 4, at noon**"
author: Zeshi Feng
output: pdf_document
fontsize: 12pt
---

```{r setup, include=FALSE}
# Some global options
knitr::opts_chunk$set(include = TRUE) # print out all code  
knitr::opts_chunk$set(echo = TRUE) # print out all output
```

**Note:**

-   Ensure that your code is fully visible in the PDF and not cropped. If needed, break the code into multiple lines to fit.

-   It is recommended to write descriptive answers outside of R code chunks (i.e., as text in the main body), while comments within the code chunks can be reserved for brief code annotations.

-   In all homework questions, include a written explanation of any output to earn full credit. For example, explain what the histograms show and how they relate to the data.

## Question 1 [20 points]

Install and load the R package `usdata`. Print the first six rows of the dataset `county_2019`.

Note: Install R packages in the Console using `install.packages("")` - this command cannot be run in the code chunks in an Rmd file. You only need to install a package once. However, you must load the package with the `library()` command each time you use it. Unlike `install.packages("")`, `library()` can be run in the code chunks in an Rmd file.

**Answer:**

```{r}
library(usdata)
head(usdata::county_2019, 6)
```


## Question 2 [20 points]

Simulate 50 random numbers from a standard normal distribution and assign them to `x1`. Simulate another 50 random numbers from a standard uniform distribution and assign them to `x2`. Calculate the following quantities for each sequence:

-   Sum
-   Minimum
-   Maximum
-   Median

Hint: You can use functions - `rnorm, runif, sum, min, max, median`

**Answer:**

```{r}
x1<-rnorm(50)
x2<-runif(50)
head(x1)
sum(x1)
min(x1)
max(x1)
median(x1)

head(x2)
sum(x2)
min(x2)
max(x2)
median(x2)

```

## Question 3 [20 points]

Visually assess the normality of `x1` and `x2` in the previous question (e.g. plot histogram).

**Answer:**

```{r}

qqnorm(x1, main = "Q-Q Plot of x1")
hist(x1, main = "Histogram of x1")
qqnorm(x2, main = "Q-Q Plot of x2")
hist(x2, main = "Histogram of x2")
```

## Question 4 [20 points]

Use the `seq` function to define a new variable, `z`, which contains 50 numbers between -5 and 5, and round these numbers to 3 decimal points. (hint: you can use functions - `seq, round`)

**Answer:**

```{r}
z <- seq(-5,5,length=50)
z <- round(z,3)
z
```

## Question 5 [20 points]

Create a scatterplot with `x1` on the horizontal axis and `z` on the vertical axis.

**Answer:**

```{r}
plot(x1,z,main="scatterplot of x1 and z",
     xlab="x1",ylab="z")
```
