---
title: "STAT 3675Q Homework 2"
subtitle: "Due date: **Thursday, September 11, at noon**"
output: pdf_document
fontsize: 12pt
author: "Zeshi Feng"
---

```{r setup, include=FALSE}
# Some global options
knitr::opts_chunk$set(include = TRUE) # print out all code  
knitr::opts_chunk$set(echo = TRUE) # print out all output
```

**Note:**

- Ensure that your code is fully visible in the PDF and not cropped. If needed, break the code into multiple lines to fit.

- It is recommended to write descriptive answers outside of R code chunks (i.e., as text in the main body), while comments within the code chunks can be reserved for brief code annotations.

- In all homework questions, include a written explanation of any output to earn full credit.


## Question 1 [25 points] 
Consider the dataset `county_2019` in the R package `usdata`. To read the documentation of the dataset, run `?county_2019`.

a. Load the dataset into the working environment as an object. Use an R function to verify that it is a data frame.

**Answer:**

```{r}
library(usdata)
data("county_2019")
class(county_2019)
```
"""library() loads the package; data() loads the dataset into your working environment; class() loads the class"""

b. Use an R function to check the numbers of observations and variables. 

**Answer:**
```{r}
dim(county_2019)
```
"""use dim() function which returns a vector of two numbers: rows, columns."""

c. Visually check the data frame. What data types does it include? Are there missing values?

**Answer:**
```{r}
str(county_2019)
anyNA(county_2019)
```
"""We use str() function to tells you the data types of each variable; Function anyNA() gives out wheather or not the data set contains NaN """

## Question 2 [25 points]

a. Generate the following sequence  using `seq()` with the `by` argument: 2 5 8 11 14 17.  Create the vector and print it out using two separate commands.

**Answer:**
```{r}
sequence_in_partA <- seq(from = 2, to = 17, by = 3)
sequence_in_partA
```
"""from, to	:the starting and (maximal) end values of the sequence. Of length 1 unless just from is supplied as an unnamed argument; by:number: increment of the sequence."

b. Generate a sequence of 15 numbers that are equally spaced between -1 and 1 without defining the increment. Hint: Read the documentation of the `seq` function. Create the vector and print it out using a single-line command.

**Answer:**
```{r}
sequence_in_partB <- seq(from = -1, to = 1, length = 15)
sequence_in_partB
```
"""same as last one, length.out: desired length of the sequence. A non-negative number, which for seq and seq.int will be rounded up if fractional."""

c. Print out the following sequences using `rep()` and `c()`, without creating any object in the environment:

i). 5 12 13 5 12 13 5 12 13

ii). 5 5 5 5 5 12 12 12

**Answer:**
```{r}
#i)
rep(c(5, 12, 13), times = 3)
#ii)
c(rep(5, 5), rep(12, 3))
```
"""nothing new, just tricks of function rep() and c(); BTW c() is a generic function which combines its arguments."""

d. Create a vector containing the 1st, 3rd, 5th, 7th, 9th and 11th elements of the sequence in part b. Print it out. 

**Answer:**
```{r}
sequence_in_partB[c(1, 3, 5, 7, 9, 11)]
```
"""Using the combined value as position of the array."""

e. What happens when a vector is multiplied by a number, a vector of the same length, or a vector of a different length? Use the vector from part a as an example. Try multiplying it by 2, `rep(2,6)`, and `c(1,2)`, respectively, using the operator `*`. Explain the results you get.

**Answer:**
```{r}
#1)
sequence_in_partA * 2
#2)
sequence_in_partA * rep(2,6)
#3)
sequence_in_partA * c(1, 2)
```
"""a) scalar multiplication; b) element-wise multiplication; c) repeats the shorter vector until lengths match"""

## Question 3 [25 points]

a. Reconsider the vector in Question 2b. Convert it to a $3\times 5$ matrix, filling the entries by column. 

**Answer:**
```{r}
my_matrix_a <- matrix(sequence_in_partB, nrow = 3, ncol = 5)
my_matrix_a
```
"""matrix(x, nrow, ncol) reshapes a vector into a matrix.
By default, R fills by column unless you add byrow = TRUE""

b. Redo part a, filling the entries by row.

**Answer:**
```{r}
my_matrix_b <- matrix(sequence_in_partB, nrow = 3, ncol = 5, byrow = TRUE)
my_matrix_b
```

c. Explain why it is unnecessary to specify both `nrow` and `ncol` for parts a and b. Then, try the following command: `matrix(c(1,2,3), nrow=5, ncol=6)`, and explain why in this case  both `nrow` and `ncol` are needed.

**Answer:**
"""If vector length is a perfect multiple of rows or columns, you only need one of nrow/ncol. If recycling is required, R cannot infer the shape automatically, so you must provide both."""

d. Extract the odd-numbered columns from the matrix created in part a. 

**Answer:**
```{r}
my_matrix_a[ , c(1, 3, 5)]
```
"""""The first blank means “all rows, c(1,3,5) means column”"

## Question 4 [25 points]

a. Explore the `set.seed()` function in R. Explain the difference between the command

```{r}
rnorm(10)
```

and the commands

```{r}
set.seed(1)
rnorm(10)
```

in terms of reproducibility.

**Answer:** 
"""set.seed(k) ensures reproducibility: same seed means same “random” results; no seed means different results"""

b. Set the random seed to 1. Create a $5\times 3$ matrix, with the $i$th column filled by randomly generated numbers from the student $t$ distribution with $i$ degrees of freedom, for $i=1,2,3$. The row names of the matrix are "1", ..., "5", and the column names are "t1", "t2", "t3". Print out the matrix.

**Answer:**
```{r}
set.seed(1)
col1 <- rt(5, df = 1)
col2 <- rt(5, df = 2)
col3 <- rt(5, df = 3) 
my_mat <- cbind(col1, col2, col3)
rownames(my_mat) <- as.character(1:5)
colnames(my_mat) <- c("t1", "t2", "t3")
my_mat
```
"""rt(n, df) generates n random numbers from a Student-t distribution with df degrees of freedom.
cbind() binds columns together."""

c. Convert the matrix to a data frame.

**Answer:**
```{r}
my_df <- as.data.frame(my_mat)
my_df
```
"""as.data.frame() takes the matrix and converts it into a data frame"""
